---
title: "Intro R"
author: "Edzer Pebesma"
date: "10/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(sf)
nc = system.file("gpkg/nc.gpkg", package = "sf") |>
    read_sf()
nc.32119 <- st_transform(nc, 'EPSG:32119')
nc.32119 |>
    select(BIR74) |>
    plot()
```
## R objects

"Everything that is, is an object"

Assignment: these two are equivalent
```{r}
a = 3
a <- 3
```
Vectors: logical, numeric (integer/double), character, raw, list 

```{r}
a = 3L
typeof(a)
a = 3.0
typeof(a)
a = 3
typeof(a)
(a = 1:3)
typeof(a)
```

### attributes

```{r}
a = 2
attr(a, "foo") = "bar"
a
```

short form:

```{r}
(a = structure(2, foo = "bar"))
attributes(a) # list all
attr(a, "foo")
attr(a, "bar")
```

### special attributes: names, dim, class

```{r}
a = 1:3
names(a) = c("a", "b", "c")
a
c(a = 1, b = 2, c = 3)
```

```{r}

```

### composite types: `factor`, `POSIXct`, `Date`, `difftime`, `data.frame`

```{r}
a = c("a", "a", "b", "c", "a", "a")
a
b = factor(a)
```
* all "scalar" objects are really vectors of length one. Almost all functions work on vectors.
* recycle rule: in

```{r}
a = 1:3
a + 1
```
the vector `1` is recycled to the length of `a`. There are catches:
```{r}
a = 1:4
a + 1:2
```



## Environments, Global Environment

By default, objects created are stored in the global environment, which is queried by
```{r}
ls()
```

## R functions

"Everything that does something, is a function call"

```{r}
3 + 4
```
is equivalent to 
```{r}
`+`(3, 4)
```
and we can even redefine
```{r}
`+`  = `-`
3 + 4
```
and remove this
```{r}
rm(`+`)
```

### pipes

## Spatial objects

```{r}
library(sf)
```

```{r}
s = st_sf(a = 1:3, b = c("a", "x", "y"), 
      geom = st_sfc(st_point(c(1,2)), st_point(c(3,2)), st_point(c(4,0))),
      crs = 'EPSG:4326')
s
attributes(s)
```

## Tidyverse

